### 2010-02-26 papers/user2010/abstract.tex
initial version of useR abstract
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@293 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[11pt]{article}
\usepackage{url}
\usepackage{vmargin}
\setpapersize{USletter}
\setmarginsrb{1in}{1in}{1in}{1in}{0pt}{0mm}{0pt}{0mm}
\usepackage{charter}

\newcommand{\proglang}[1]{\textsf{#1}}
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}

\author{
Romain Fran\c{c}ois\\ {\small \url{romain@r-enthusiasts.com} } \and
Dirk Eddelbuettel\\ {\small \url{edd@debian.org} }
}

\title{\pkg{RProtoBuf}: Protocol Buffers for R }
\date{Abstract submitted for \textsl{useR! 2010}}

\begin{document}

\maketitle
\thispagestyle{empty}
\begin{abstract}
  \addtolength{\parskip}{\baselineskip} 	% add a little vertical space
  \noindent % no ident for first paragraph
  Protocol buffers are a flexible, efficient, 
  automated mechanism for serializing structured data---think XML, but 
  smaller, faster, and simpler. 
  You define how you want your data to be structured once, then you can use 
  special generated source code to easily write and read your structured 
  data to and from a variety of data streams and using a 
  variety of officially supported languages---
  \proglang{Java}, \proglang{C++}, or \proglang{Python}---or 
  third party implementations for languages such as \proglang{C\#}, 
  \proglang{Perl}, \proglang{Ruby}, \proglang{Visual Basic}, and now 
  \proglang{R} via the \pkg{RProtoBuf} package.

  The \pkg{RProtoBuf} package implements R bindings to the C++
  protobuf library from Google. It uses features of the protocol 
  buffer library to support creation, manipulation, parsing and serialization
  of protocol buffers messages. The \pkg{RProtoBuf} package uses S4 classes
  and external pointers to expose objects that look and feel like standard 
  R lists, but constrained by the language agnostic definition of the 
  message type.

  \noindent \textbf{Keywords:}  
  \proglang{R}, \proglang{C++}, Serialization, Data Interchange
\end{abstract}
\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: