### 2013-12-26 papers/rjournal/eddelbuettel-francois-stokely.Rnw
Document file descriptors and enumvalue descriptors, and add a summary table showing the number of slots, methods, and dynamically dispatched fields of each class.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@594 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

% !TeX root = RJwrapper.tex
% We don't want a left margin for Sinput or Soutput for our table 1.
%\DefineVerbatimEnvironment{Sinput}{Verbatim} {xleftmargin=0em}
%\DefineVerbatimEnvironment{Soutput}{Verbatim}{xleftmargin=0em}
%\DefineVerbatimEnvironment{Scode}{Verbatim}{xleftmargin=2em}
% Setting the topsep to 0 reduces spacing from input to output and
% improves table 1.
\fvset{listparameters={\setlength{\topsep}{0pt}}}
\renewenvironment{Schunk}{\vspace{\topsep}}{\vspace{\topsep}}

\title{RProtoBuf: Efficient Cross-Language Data Serialization in R}
\author{by Dirk Eddelbuettel, Romain Fran\c{c}ois, and Murray Stokely}

\maketitle

\abstract{Modern data collection and analysis pipelines often involve
 a sophisticated mix of applications written in general purpose and
 specialized programming languages.  Protocol Buffers are a popular
 method of serializing structured data between applications---while remaining
 independent of programming languages or operating system.  The
 \CRANpkg{RProtoBuf} package provides a complete interface to this
 library.
 %TODO(ms) keep it less than 150 words.
}

%TODO(de) 'protocol buffers' or 'Protocol Buffers' ?

\section{Introduction}

Modern data collection and analysis pipelines are increasingly being
built using collections of components to better manage software
complexity through reusability, modularity, and fault
isolation \citep{Wegiel:2010:CTT:1932682.1869479}.  Different
programming languages are often used for the different phases of data
analysis -- collection, cleaning, analysis, post-processing, and
presentation in order to take advantage of the unique combination of 
performance, speed of development, and library support offered by
different environments.  Each stage of the data
analysis pipeline may involve storing intermediate results in a
file or sending them over the network.  Programming langauges such as
Java, Ruby, Python, and R include built-in serialization support, but
these formats are tied to the specific programming language in use.
% TODO(ms): and they often don't support versioning among other faults.
CSV files can be read and written by many applications and so are
often used for exporting tabular data.  However, CSV files have a
number of disadvantages, such as a limitation of exporting only
tabular datasets, lack of type-safety, inefficient text representation
and parsing, and abiguities in the format involving special
characters.  JSON is another widely supported format used mostly on
the web that removes many of these disadvantages, but it too suffers
from being too slow to parse and also does not provide strong typing
between integers and floating point.  Large numbers of JSON messages
would also be required to duplicate the field names with each message.

TODO(ms): Also work in reference to Split-Apply-Combine pattern for
data analysis \citep{wickham2011split}, since that is a great pattern
but it seems overly optimistic to expect all of those phases to always
be done in the same language.

This article describes the basics of Google's Protocol Buffers through
an easy to use R package, \CRANpkg{RProtoBuf}.  After describing the
basics of protocol buffers and \CRANpkg{RProtoBuf}, we illustrate
several common use cases for protocol buffers in data analysis.

\section{Protocol Buffers}

Once the data serialization needs get complex enough, application
developers typically benefit from the use of an \emph{interface
description language}, or \emph{IDL}.  IDLs like Google's Protocol
Buffers and Apache Thrift provide a compact well-documented schema for
cross-langauge data structures as well efficient binary interchange
formats.  The schema can be used to generate model classes for
statically typed programming languages such as C++ and Java, or can be
used with reflection for dynamically typed programming languages.
Since the schema is provided separately from the encoded data, the
data can be efficiently encoded to minimize storage costs of the
stored data when compared with simple ``schema-less'' binary
interchange formats like BSON.

% TODO(ms) Also talk about versioning and why its useful.

%BSON, msgpack, Thrift, and Protocol Buffers take this latter approach,
%with the

% There are references comparing these we should use here.

TODO Also mention Thrift and msgpack and the references comparing some
of these tradeoffs.

Introductory section which may include references in parentheses
\citep{R}, or cite a reference such as \citet{R} in the text.

Protocol buffers are a language-neutral, platform-neutral, extensible
way of serializing structured data for use in communications
protocols, data storage, and more.

Protocol Buffers offer key features such as an efficient data interchange
format that is both language- and operating system-agnostic yet uses a
lightweight and highly performant encoding, object serialization and
de-serialization as well data and configuration management. Protocol
buffers are also forward compatible: updates to the \texttt{proto}
files do not break programs built against the previous specification.

While benchmarks are not available, Google states on the project page that in
comparison to XML, protocol buffers are at the same time \textsl{simpler},
between three to ten times \textsl{smaller}, between twenty and one hundred
times \textsl{faster}, as well as less ambiguous and easier to program.

The protocol buffers code is released under an open-source (BSD) license. The
protocol buffer project (\url{http://code.google.com/p/protobuf/})
contains a C++ library and a set of runtime libraries and compilers for
C++, Java and Python.

With these languages, the workflow follows standard practice of so-called
Interface Description Languages (IDL)
(c.f. \href{http://en.wikipedia.org/wiki/Interface_description_language}{Wikipedia
  on IDL}).  This consists of compiling a protocol buffer description file
(ending in \texttt{.proto}) into language specific classes that can be used
to create, read, write and manipulate protocol buffer messages. In other
words, given the 'proto' description file, code is automatically generated
for the chosen target language(s). The project page contains a tutorial for
each of these officially supported languages:
\url{http://code.google.com/apis/protocolbuffers/docs/tutorials.html}

Besides the officially supported C++, Java and Python implementations, several projects have been
created to support protocol buffers for many languages. The list of known
languages to support protocol buffers is compiled as part of the
project page: \url{http://code.google.com/p/protobuf/wiki/ThirdPartyAddOns}

\section{Basic Usage: Messages and Descriptors}

This section describes how to use the R API to create and manipulate
protocol buffer messages in R, and how to read and write the
binary \emph{payload} of the messages to files and arbitrary binary
R connections.

The two fundamental building blocks of Protocol Buffers are Messages
and Descriptors.  Messages provide a common abstract encapsulation of
structured data fields of the type specified in a Message Descriptor.
Message Descriptors are defined in \texttt{.proto} files and define a
schema for a particular named class of messages.  This separation
between schema and the message objects is in contrast to
more verbose formats like JSON, and when combined with the efficient
binary representation of any Message object explains a large part of
the performance and storage-space advantage offered by Protocol
Buffers. TODO(ms): we already said some of this above.  clean up.

Table~\ref{tab:proto} shows an example \texttt{.proto} file which
defines the \texttt{tutorial.Person} type.  The R code in the right
column shows an example of creating a new message of this type and
populating its fields.

% lifted from protobuf page:
%With Protocol Buffers you define how you want your data to be
%structured once, and then you can read or write structured data to and
%from a variety of data streams using a variety of different
%languages.  The definition

\noindent
\begin{table}
\begin{tabular}{@{\hskip .01\textwidth}p{.40\textwidth}@{\hskip .015\textwidth}|@{\hskip .015\textwidth}p{0.55\textwidth}@{\hskip .01\textwidth}}
\hline
Schema : \texttt{addressbook.proto} & Example R Session\\
\hline
\begin{minipage}{.35\textwidth}
\vspace{2mm}
\begin{example}
package tutorial;
message Person {
 required string name = 1;
 required int32 id = 2;
 optional string email = 3;
 enum PhoneType {
   MOBILE = 0; HOME = 1;
   WORK = 2;
 }
 message PhoneNumber {
   required string number = 1;
   optional PhoneType type = 2;
 }
 repeated PhoneNumber phone = 4;
}
\end{example}
\vspace{2mm}
\end{minipage} & \begin{minipage}{.45\textwidth}
<<echo=TRUE>>=
library(RProtoBuf)
person <- new(tutorial.Person, id=1, name="Dirk")
person
person$name
person$name <- "Romain"
cat(as.character(person))
serialize(person, NULL)
@ 
\end{minipage} \\
\hline
\end{tabular}
\caption{The schema representation from a \texttt{.proto} file for the
  \texttt{tutorial.Person} class (left) and simple R code for creating
  an object of this class and accessing its fields (right).}
\label{tab:proto}
\end{table}

%This section may contain a figure such as Figure~\ref{figure:rlogo}.
%
%\begin{figure}[htbp]
%  \centering
%  \includegraphics{Rlogo}
%  \caption{The logo of R.}
%  \label{figure:rlogo}
%\end{figure}

\subsection{Importing Message Descriptors from \texttt{.proto} files}

%The three basic abstractions of \CRANpkg{RProtoBuf} are Messages,
%which encapsulate a data structure, Descriptors, which define the
%schema used by one or more messages, and DescriptorPools, which
%provide access to descriptors.

Before we can create a new Protocol Buffer Message or parse a
serialized stream of bytes as a Message, we must read in the message
type specification from a \texttt{.proto} file.

New \texttt{.proto} files are imported with the \code{readProtoFiles}
function, which can import a single file, all files in a directory, or
all \texttt{.proto} files provided by another R package.

The \texttt{.proto} file syntax for defining the structure of protocol
buffer data is described comprehensively on Google Code:
\url{http://code.google.com/apis/protocolbuffers/docs/proto.html}.

Once the proto files are imported, all message descriptors are
are available in the R search path in the \texttt{RProtoBuf:DescriptorPool}
special environment. The underlying mechanism used here is
described in more detail in Section~\ref{sec-lookup}.

<<>>=
ls( "RProtoBuf:DescriptorPool" )
@

%\subsection{Importing proto files}
%In contrast to the other languages (Java, C++, Python) that are officially
%supported by Google, the implementation used by the \texttt{RProtoBuf}
%package does not rely on the \texttt{protoc} compiler (with the exception of
%the two functions discussed in the previous section). This means that no
%initial step of statically compiling the proto file into C++ code that is
%then accessed by R code is necessary. Instead, \texttt{proto} files are
%parsed and processed \textsl{at runtime} by the protobuf C++ library---which
%is much more appropriate for a dynamic language.

\subsection{Creating a message}

New messages are created with the \texttt{new} function which accepts
a Message Descriptor and optionally a list of ``name = value'' pairs
to set in the message.
%The objects contained in the special environment are
%descriptors for their associated message types. Descriptors will be
%discussed in detail in another part of this document, but for the
%purpose of this section, descriptors are just used with the \texttt{new}
%function to create messages.

<<>>=
p1 <- new( tutorial.Person )
p <- new( tutorial.Person, name = "Romain", id = 1 )
@

\subsection{Access and modify fields of a message}

Once the message is created, its fields can be queried
and modified using the dollar operator of R, making protocol
buffer messages seem like lists.

<<>>=
p$name
p$id
p$email <- "francoisromain@free.fr"
@

However, as opposed to R lists, no partial matching is performed
and the name must be given entirely.

The \verb|[[| operator can also be used to query and set fields
of a mesages, supplying either their name or their tag number :

<<>>=
p[["name"]] <- "Romain Francois"
p[[ 2 ]] <- 3
p[[ "email" ]]
@

Protocol buffers include a 64-bit integer type, but R lacks native
64-bit integer support.  A workaround is available and described in
Section~\ref{sec:int64} for working with large integer values.

% TODO(mstokely): Document extensions here.
% There are none in addressbook.proto though.

\subsection{Display messages}

Protocol buffer messages and descriptors implement \texttt{show}
methods that provide basic information about the message :

<<>>=
p
@

For additional information, such as for debugging purposes,
the \texttt{as.character} method provides a more complete ASCII
representation of the contents of a message.

<<>>=
writeLines( as.character( p ) )
@

\subsection{Serializing messages}

However, the main focus of protocol buffer messages is
efficiency. Therefore, messages are transported as a sequence
of bytes. The \texttt{serialize} method is implemented for
protocol buffer messages to serialize a message into a sequence of
bytes that represents the message.
%(raw vector in R speech) that represents the message.

<<>>=
serialize( p, NULL )
@

The same method can also be used to serialize messages to files :

<<>>=
tf1 <- tempfile()
serialize( p, tf1 )
readBin( tf1, raw(0), 500 )
@

Or to arbitrary binary connections:

<<>>=
tf2 <- tempfile()
con <- file( tf2, open = "wb" )
serialize( p, con )
close( con )
readBin( tf2, raw(0), 500 )
@

\texttt{serialize} can also be used in a more traditional
object oriented fashion using the dollar operator :

<<>>=
# serialize to a file
p$serialize( tf1 )
# serialize to a binary connection
con <- file( tf2, open = "wb" )
p$serialize( con )
close( con )
@

\subsection{Parsing messages}

The \texttt{RProtoBuf} package defines the \texttt{read} and
\texttt{readASCII} functions to read messages from files, raw vectors,
or arbitrary connections.  \texttt{read} expects to read the message
payload from binary files or connections and \texttt{readASCII} parses
the human-readable ASCII output that is created with
\code{as.character}.

The binary representation of the message (often called the payload)
does not contain information that can be used to dynamically
infer the message type, so we have to provide this information
to the \texttt{read} function in the form of a descriptor :

<<>>=
msg <- read( tutorial.Person, tf1 )
writeLines( as.character( msg ) )
@

The \texttt{input} argument of \texttt{read} can also be a binary
readable R connection, such as a binary file connection:

<<>>=
con <- file( tf2, open = "rb" )
message <- read( tutorial.Person, con )
close( con )
writeLines( as.character( message ) )
@

Finally, the payload of the message can be used :

<<>>=
# reading the raw vector payload of the message
payload <- readBin( tf1, raw(0), 5000 )
message <- read( tutorial.Person, payload )
@

\texttt{read} can also be used as a pseudo method of the descriptor
object :

<<>>=
# reading from a file
message <- tutorial.Person$read( tf1 )
# reading from a binary connection
con <- file( tf2, open = "rb" )
message <- tutorial.Person$read( con )
close( con )
# read from the payload
message <- tutorial.Person$read( payload )
@

\section{Under the hood: S4 Classes, Methods, and Pseudo Methods}

The \CRANpkg{RProtoBuf} package uses the S4 system to store
information about descriptors and messages.  Using the S4 system
allows the \texttt{RProtoBuf} package to dispatch methods that are not
generic in the S3 sense, such as \texttt{new} and
\texttt{serialize}.

Each R object stores an external pointer to an object managed by
the \texttt{protobuf} C++ library.
The \CRANpkg{Rcpp} \citep{eddelbuettel2011rcpp} package is used to
facilitate the integration of the R and C++ code for these objects.

% Message, Descriptor, FieldDescriptor, EnumDescriptor,
% FileDescriptor, EnumValueDescriptor
%
% grep RPB_FUNC * | grep -v define|wc -l
% 84
% grep RPB_ * | grep -v RPB_FUNCTION | grep METHOD|wc -l
% 33

There are over 100 C++ functions that provide the glue code between
the member functions of the 6 primary Message and Descriptor classes
in the protobuf library.  Wrapping each method individually allows us
to add user friendly custom error handling, type coercion, and
performance improvements at the cost of a more verbose
implementation.  The RProtoBuf implementation in many ways motivated
the development of Rcpp Modules \citep{eddelbuettel2010exposing},
which provide a more concise way of wrapping C++ functions and classes
in a single entity.

The \texttt{RProtoBuf} package combines the \emph{R typical} dispatch
of the form \verb|method( object, arguments)| and the more traditional
object oriented notation \verb|object$method(arguments)|.
Additionally, \texttt{RProtoBuf} implements the \texttt{.DollarNames} S3 generic function
(defined in the \texttt{utils} package) for all classes to enable tab
completion.  Completion possibilities include pseudo method names for all
classes, plus dynamic dispatch on names or types specific to a given object.

% TODO(ms): Add column check box for doing dynamic dispatch based on type.
\begin{table}[h]
\centering
\begin{tabular}{|l|c|c|l|}
\hline
\textbf{Class} & \textbf{Slots} & \textbf{Methods} & \textbf{Dynamic Dispatch}\\
\hline
\hline
Message & 2 & 20 & yes (field names)\\
\hline
Descriptor & 2 & 16 & yes (field names, enum types, nested types)\\
\hline
FieldDescriptor & 4 & 18 & no\\
\hline
EnumDescriptor & 4 & 11 & yes (enum constant names)\\
\hline
FileDescriptor & 3 & 6 & yes (message/field definitions)\\
\hline
EnumValueDescriptor & 3 & 6 & no\\
\hline
\end{tabular}
\end{table}

\subsection{Messages}

The \texttt{Message} S4 class represents Protocol Buffer Messages and
is the core abstraction of \CRANpkg{RProtoBuf}.  The class contains
the slots \texttt{pointer} and \texttt{type} as described on the
Table~\ref{Message-class-table}.

\begin{table}[h]
\centering
\begin{tabular}{|cp{10cm}|}
\hline
\textbf{Slot} & \textbf{Description} \\
\hline
\texttt{pointer} & External pointer to the \texttt{Message} object of the C++ proto library. Documentation for the
\texttt{Message} class is available from the protocol buffer project page:
\url{http://code.google.com/apis/protocolbuffers/docs/reference/cpp/google.protobuf.message.html#Message} \\
\hline
\texttt{type} & Fully qualified name of the message. For example a \texttt{Person} message
has its \texttt{type} slot set to \texttt{tutorial.Person} \\
\hline
\end{tabular}
\caption{\label{Message-class-table}Description of slots for the \texttt{Message} S4 class}
\end{table}

Each \texttt{Message} contains a pointer to a \texttt{Descriptor}
which defines the schema of the data defined in the Message, as well
as a number of \texttt{FieldDescriptors} for the individual fields of
the message.  In addition to the field name extractors of
\texttt{Messages} introduced in the previous section, a complete list
of Message methods is available in Table~\ref{Message-methods-table}.

\begin{table}[h]
\centering
\begin{small}
\begin{tabular}{l|l}
\textbf{Method} & \textbf{Description} \\
\hline
\hline
\texttt{has} & Indicates if a message has a given field.   \\
\texttt{clone} & Creates a clone of the message \\
\texttt{isInitialized} & Indicates if a message has all its required fields set\\
\texttt{serialize} & serialize a message to a file, binary connection, or raw vector\\
\texttt{clear} & Clear one or several fields of a message, or the entire message\\
\texttt{size} & The number of elements in a message field\\
\texttt{bytesize} & The number of bytes the message would take once serialized\\
\hline
\texttt{swap} & swap elements of a repeated field of a message\\
\texttt{set} & set elements of a repeated field\\
\texttt{fetch} & fetch elements of a repeated field\\
\texttt{setExtension} & set an extension of a message\\
\texttt{getExtension} & get the value of an extension of a message\\
\texttt{add} & add elements to a repeated field \\
\hline
\texttt{str} & the R structure of the message\\
\texttt{as.character} & character representation of a message\\
\texttt{toString} & character representation of a message (same as \texttt{as.character}) \\
\texttt{as.list} & converts message to a named R list\\
\texttt{update} & updates several fields of a message at once\\
\texttt{descriptor} & get the descriptor of the message type of this message\\
\texttt{fileDescriptor} & get the file descriptor of this message's descriptor\\
\hline
\end{tabular}
\end{small}
\caption{\label{Message-methods-table}Description of methods for the \texttt{Message} S4 class}
\end{table}

\subsection{Descriptors}

Message descriptors are represented in R with the
\emph{Descriptor} S4 class. The class contains
the slots \texttt{pointer} and \texttt{type} :

\begin{table}[h]
\centering
\begin{tabular}{|cp{10cm}|}
\hline
\textbf{Slot} & \textbf{Description} \\
\hline
\texttt{pointer} & External pointer to the \texttt{Descriptor} object of the C++ proto library. Documentation for the
\texttt{Descriptor} class is available from the protocol buffer project page:
\url{http://code.google.com/apis/protocolbuffers/docs/reference/cpp/google.protobuf.descriptor.html#Descriptor} \\
\hline
\texttt{type} & Fully qualified path of the message type. \\
\hline
\end{tabular}
\caption{\label{Descriptor-class-table}Description of slots for the \texttt{Descriptor} S4 class}
\end{table}

Similarly to messages, the \verb|$| operator can be used to retrieve
descriptors that are contained in the descriptor, or invoke
pseudo-methods.  Thise can be used to extract field descriptors, enum
descriptors, or descriptors for a nested type.

<<>>=
# field descriptor
tutorial.Person$email

# enum descriptor
tutorial.Person$PhoneType

# nested type descriptor
tutorial.Person$PhoneNumber
# same as
tutorial.Person.PhoneNumber
@

Table~\ref{Descriptor-methods-table} provides a complete list of the
avalailable methods for Descriptors.

\begin{table}[h]
\centering
\begin{small}
\begin{tabular}{l|l}
\textbf{Method} & \textbf{Description} \\
\hline
\hline
\texttt{new} & Creates a prototype of a message described by this descriptor.\\
\texttt{read} & Reads a message from a file or binary connection.\\
\texttt{readASCII} & Read a message in ASCII format from a file or
text connection.\\
\hline
\texttt{name} & Retrieve the name of the message type associated with
this descriptor.\\
\texttt{as.character} & character representation of a descriptor\\
\texttt{toString} & character representation of a descriptor (same as \texttt{as.character}) \\
\texttt{as.list} & return a named
list of the field, enum, and nested descriptors included in this descriptor.\\
\texttt{asMessage} & return DescriptorProto message. \\
\hline
\texttt{fileDescriptor} & Retrieve the file descriptor of this
descriptor.\\
\texttt{containing\_type} & Retrieve the descriptor describing the message type containing this descriptor.\\
\texttt{field\_count} & Return the number of fields in this descriptor.\\
\texttt{field} & Return the descriptor for the specified field in this descriptor.\\
\texttt{nested\_type\_count} & The number of nested types in this descriptor.\\
\texttt{nested\_type} & Return the descriptor for the specified nested
type in this descriptor.\\
\texttt{enum\_type\_count} & The number of enum types in this descriptor.\\
\texttt{enum\_type} & Return the descriptor for the specified enum
type in this descriptor.\\
\hline
\end{tabular}
\end{small}
\caption{\label{Descriptor-methods-table}Description of methods for the \texttt{Descriptor} S4 class}
\end{table}

\subsection{field descriptors}
\label{subsec-field-descriptor}

The class \emph{FieldDescriptor} represents field
descriptor in R. This is a wrapper S4 class around the
\texttt{google::protobuf::FieldDescriptor} C++ class.
Table~\ref{fielddescriptor-methods-table} describes the methods
defined for the \texttt{FieldDescriptor} class.

\begin{table}[h]
\centering
\begin{tabular}{|cp{10cm}|}
\hline
\textbf{Slot} & \textbf{Description} \\
\hline
\texttt{pointer} & External pointer to the \texttt{FieldDescriptor} C++ variable \\
\hline
\texttt{name} & Simple name of the field \\
\hline
\texttt{full\_name} & Fully qualified name of the field \\
\hline
\texttt{type} & Name of the message type where the field is declared \\
\hline
\end{tabular}
\caption{\label{FieldDescriptor-class-table}Description of slots for the \texttt{FieldDescriptor} S4 class}
\end{table}

\begin{table}[h]
\centering
\begin{small}
\begin{tabular}{l|l}
\hline
\textbf{Method} & \textbf{Description} \\
\hline
\hline
\texttt{as.character} & Character representation of a descriptor\\
\texttt{toString} & Character
representation of a descriptor (same as \texttt{as.character}) \\
\texttt{asMessage} & Return FieldDescriptorProto message. \\
\texttt{name} & Return the name of the field descriptor.\\
\texttt{fileDescriptor} & Return the fileDescriptor where this field is defined.\\
\texttt{containing\_type} & Return the containing descriptor of this field.\\
\texttt{is\_extension} & Return TRUE if this field is an extension.\\
\texttt{number} & Gets the declared tag number of the field.\\
\texttt{type} & Gets the type of the field.\\
\texttt{cpp\_type} & Gets the C++ type of the field.\\
\texttt{label} & Gets the label of a field (optional, required, or repeated).\\
\texttt{is\_repeated} & Return TRUE if this field is repeated.\\
\texttt{is\_required} & Return TRUE if this field is required.\\
\texttt{is\_optional} & Return TRUE if this field is optional.\\
\texttt{has\_default\_value} & Return TRUE if this field has a default value.\\
\texttt{default\_value} & Return the default value.\\
\texttt{message\_type} & Return the message type if this is a message type field.\\
\texttt{enum\_type} & Return the enum type if this is an enum type field.\\
\hline
\end{tabular}
\end{small}
\caption{\label{fielddescriptor-methods-table}Description of methods for the \texttt{FieldDescriptor} S4 class}
\end{table}

% TODO(ms): Useful distinction to make -- FieldDescriptor does not do
% separate '$' dispatch like Messages, Descriptors, and
% EnumDescriptors do.  Should it?

\subsection{enum descriptors}
\label{subsec-enum-descriptor}

The class \emph{EnumDescriptor} is an R wrapper
class around the C++ class \texttt{google::protobuf::EnumDescriptor}.
Table~\ref{enumdescriptor-methods-table} describes the methods
defined for the \texttt{EnumDescriptor} class.

The \verb|$| operator can be used to retrieve the value of enum
constants contained in the EnumDescriptor, or to invoke
pseudo-methods.

<<>>=
tutorial.Person$PhoneType
tutorial.Person$PhoneType$WORK
@

\begin{table}[h]
\centering
\begin{tabular}{|cp{10cm}|}
\hline
\textbf{Slot} & \textbf{Description} \\
\hline
\texttt{pointer} & External pointer to the \texttt{EnumDescriptor} C++ variable \\
\hline
\texttt{name} & Simple name of the enum \\
\hline
\texttt{full\_name} & Fully qualified name of the enum \\
\hline
\texttt{type} & Name of the message type where the enum is declared \\
\hline
\end{tabular}
\caption{\label{EnumDescriptor-class-table}Description of slots for the \texttt{EnumDescriptor} S4 class}
\end{table}

\begin{table}[h]
\centering
\begin{small}
\begin{tabular}{l|l}
\hline
\textbf{Method} & \textbf{Description} \\
\hline
\hline
\texttt{as.list} & return a named
integer vector with the values of the enum and their names.\\
\texttt{as.character} & character representation of a descriptor\\
\texttt{toString} & character
representation of a descriptor (same as \texttt{as.character}) \\
\texttt{asMessage} & return EnumDescriptorProto message. \\
\texttt{name} & Return the name of the enum descriptor.\\
\texttt{fileDescriptor} & Return the fileDescriptor where this field is defined.\\
\texttt{containing\_type} & Return the containing descriptor of this field.\\
\texttt{length} & Return the number of constants in this enum.\\
\texttt{has} & Return TRUE if this enum contains the specified named constant string.\\
\texttt{value\_count} & Return the number of constants in this enum (same as \texttt{length}).\\
\texttt{value} & Return the EnumValueDescriptor of an enum value of specified index, name, or number.\\
\hline
\end{tabular}
\end{small}
\caption{\label{enumdescriptor-methods-table}Description of methods for the \texttt{EnumDescriptor} S4 class}
\end{table}

\subsection{file descriptors}
\label{subsec-file-descriptor}

The class \emph{FileDescriptor} is an R wrapper
class around the C++ class \texttt{google::protobuf::FileDescriptor}.
Table~\ref{filedescriptor-methods-table} describes the methods
defined for the \texttt{FileDescriptor} class.

The \verb|$| operator can be used to retrieve named fields defined in
the FileDescriptor, or to invoke pseudo-methods.

<<>>=
f <- tutorial.Person$fileDescriptor()
f
f$Person
@

\begin{table}[h]
\centering
\begin{tabular}{|cp{10cm}|}
\hline
\textbf{slot} & \textbf{description} \\
\hline
\texttt{pointer} & external pointer to the \texttt{FileDescriptor} object of the C++ proto library. Documentation for the
\texttt{FileDescriptor} class is available from the protocol buffer project page:
\url{http://developers.google.com/protocol-buffers/docs/reference/cpp/google.protobuf.descriptor.html#FileDescriptor} \\
\hline
\texttt{filename} & fully qualified pathname of the \texttt{.proto} file.\\
\hline
\texttt{package} & package name defined in this \texttt{.proto} file.\\
\hline
\end{tabular}
\caption{\label{FileDescriptor-class-table}Description of slots for the \texttt{FileDescriptor} S4 class}
\end{table}

\begin{table}[h]
\centering
\begin{small}
\begin{tabular}{l|l}
\hline
\textbf{Method} & \textbf{Description} \\
\hline
\hline
\texttt{name} & Return the filename for this FileDescriptorProto.\\
\texttt{package} & Return the file-level package name specified in this FileDescriptorProto.\\
\texttt{as.character} & character representation of a descriptor. \\
\texttt{toString} & character representation of a descriptor (same as \texttt{as.character}). \\
\texttt{asMessage} & return FileDescriptorProto message. \\
\texttt{as.list} & return named list of descriptors defined in this file descriptor.\\
\hline
\end{tabular}
\end{small}
\caption{\label{filedescriptor-methods-table}Description of methods for the \texttt{FileDescriptor} S4 class}
\end{table}

\subsection{enum value descriptors}
\label{subsec-enumvalue-descriptor}

The class \emph{EnumValueDescriptor} is an R wrapper
class around the C++ class \texttt{google::protobuf::EnumValueDescriptor}.
Table~\ref{EnumValueDescriptor-methods-table} describes the methods
defined for the \texttt{EnumValueDescriptor} class.

The \verb|$| operator can be used to invoke pseudo-methods.

<<>>=
tutorial.Person$PhoneType$value(1)
tutorial.Person$PhoneType$value(name="HOME")
tutorial.Person$PhoneType$value(number=1)
@

\begin{table}[h]
\centering
\begin{tabular}{|cp{10cm}|}
\hline
\textbf{slot} & \textbf{description} \\
\hline
\texttt{pointer} & External pointer to the \texttt{EnumValueDescriptor} C++ variable \\
\hline
\texttt{name} & simple name of the enum value \\
\hline
\texttt{full\_name} & fully qualified name of the enum value \\
\hline
\end{tabular}
\caption{\label{EnumValueDescriptor-class-table}Description of slots for the \texttt{EnumValueDescriptor} S4 class}
\end{table}

\begin{table}[h]
\centering
\begin{small}
\begin{tabular}{l|l}
\hline
\textbf{Method} & \textbf{Description} \\
\hline
\hline
\texttt{number} & return the number of this EnumValueDescriptor. \\
\texttt{name} & Return the name of the enum value descriptor.\\
\texttt{enum\_type} & return the EnumDescriptor type of this EnumValueDescriptor. \\
\texttt{as.character} & character representation of a descriptor. \\
\texttt{toString} & character representation of a descriptor (same as \texttt{as.character}). \\
\texttt{asMessage} & return EnumValueDescriptorProto message. \\
\hline
\end{tabular}
\end{small}
\caption{\label{enumvaluedescriptor-methods-table}Description of methods for the \texttt{EnumValueDescriptor} S4 class}
\end{table}

\section{Type Coercion}

\subsection{Booleans}
Bools
Int64s.

\subsection{64-bit integers}
\label{sec:int64}

R does not have native 64-bit integer support.  Instead, R treats
large integers as doubles which have limited precision.  For example,
it loses the ability to distinguish some distinct integers:

<<>>=
2^53 == (2^53 + 1)
@

Protocol Buffers are frequently used to pass data between different
systems, however, and most other systems these days have support for
64-bit integers.  To work around this, RProtoBuf allows users to get
and set 64-bit integer types by treating them as characters.

<<echo=FALSE,print=FALSE>>=
if (!exists("protobuf_unittest.TestAllTypes",
            "RProtoBuf:DescriptorPool")) {
    unittest.proto.file <- system.file("unitTests", "data",
                                       "unittest.proto",
                                       package="RProtoBuf")
    readProtoFiles(file=unittest.proto.file)
}
@

If we try to set an int64 field in R to double values, we lose
precision:

<<>>=
test <- new(protobuf_unittest.TestAllTypes)
test$repeated_int64 <- c(2^53, 2^53+1)
length(unique(test$repeated_int64))
@

However, we can specify the values as character strings so that the
C++ library on which RProtoBuf is based can store a true 64-bit
integer representation of the data.

<<>>=
test$repeated_int64 <- c("9007199254740992", "9007199254740993")
@

When reading the value back into R, numeric types are returned by
default, but when the full precision is required a character value
will be returned if the \texttt{RProtoBuf.int64AsString} option is set
to \texttt{TRUE}.

<<>>=
options("RProtoBuf.int64AsString" = FALSE)
test$repeated_int64
length(unique(test$repeated_int64))
options("RProtoBuf.int64AsString" = TRUE)
test$repeated_int64
length(unique(test$repeated_int64))
@

<<echo=FALSE,print=FALSE>>=
options("RProtoBuf.int64AsString" = FALSE)
@

\section{Related work on IDLs (greatly expanded from what you have)}

\section{Design tradeoffs: reflection vs proto compiler (not addressed
  at all in current vignettes)}

\subsection{Performance considerations}

TODO RProtoBuf is quite flexible and easy to use for interactive
analysis, but it is not designed for certain classes of operations one
might like to do with protocol buffers.  For example, taking a list of
10,000 protocol buffers, extracting a named field from each one, and
computing a aggregate statistics on those values would be extremely
slow with RProtoBuf, and while this is a useful class of operations,
it is outside of the scope of RProtoBuf.  We should be very clear
about this to clarify the goals and strengths of RProtoBuf and its
reflection and object mapping.

\subsection{Serialization comparison}

TODO comparison of protobuf serialization sizes/times for various vectors.  Compared to R's native serialization.  Discussion of the RHIPE approach of serializing any/all R objects, vs more specific protocol buffers for specific R objects.

\section{Descriptor lookup}
\label{sec-lookup}

The \texttt{RProtoBuf} package uses the user defined tables framework
that is defined as part of the \texttt{RObjectTables} package available
from the OmegaHat project \citep{RObjectTables}.

The feature allows \texttt{RProtoBuf} to install the
special environment \emph{RProtoBuf:DescriptorPool} in the R search path.
The environment is special in that, instead of being associated with a
static hash table, it is dynamically queried by R as part of R's usual
variable lookup. In other words, it means that when the R interpreter
looks for a binding to a symbol (foo) in its search path,
it asks to our package if it knows the binding "foo", this is then
implemented by the \texttt{RProtoBuf} package by calling an internal
method of the \texttt{protobuf} C++ library.

\section{Other approaches}

Saptarshi Guha wrote another package that deals with integration
of protocol buffer messages with R, taking a different angle :
serializing any R object as a message, based on a single catch-all
\texttt{proto} file.  Saptarshi's package is available at
\url{http://ml.stat.purdue.edu/rhipe/doc/html/ProtoBuffers.html}.

Jeroen Ooms took a similar approach influenced by Saptarshi in his
\pkg{RProtoBufUtils} package.  Unlike Saptarshi's package,
RProtoBufUtils depends on RProtoBuf for underlying message operations.
This package is available at
\url{https://github.com/jeroenooms/RProtoBufUtils}.

% Phillip Yelland wrote another implementation, currently proprietary,
% that has significant speed advantages when querying fields from a
% large number of protocol buffers, but is less user friendly for the
% basic cases documented here.

\section{Basic usage example - tutorial.Person}

\section{Application: distributed Data Collection with MapReduce}

We could describe a common MapReduce pattern of having the MR written
in another language output protocol buffers that are later pulled into
R.  There is some text about this in section 2 of
http://cran.r-project.org/web/packages/HistogramTools/vignettes/HistogramTools.pdf

\section{Application: Sending/receiving Interaction With Servers}

Unlike Apache Thrift, Protocol Buffers do not include a concrete RPC
implementation.  However, serialized protocol buffers can trivially be
sent over TCP or integrated with a proprietary RPC system.  Combined
with an RPC system this means that one can interactively craft request
messages, send the serialized message to a remote server, read back a
response, and then parse the response protocol buffer interactively.

\section{Summary}

TODO(ms): random citations to work in:

Many sources compare data serialization formats and show protocol
buffers very favorably to the alternatives, such
as \citep{Sumaray:2012:CDS:2184751.2184810}

%Its pretty useful.  Murray to see if he can get approval to talk a
%tiny bit about how much its used at Google.

%This file is only a basic article template. For full details of \emph{The R Journal} style and information on how to prepare your article for submission, see the \href{http://journal.r-project.org/latex/RJauthorguide.pdf}{Instructions for Authors}.

\bibliography{eddelbuettel-francois-stokely}

\address{Dirk Eddelbuettel\\
  Debian and R Projects\\
  711 Monroe Avenue, River Forest, IL 60305\\
  USA}
\email{edd@debian.org}

\address{Author Two\\
  Affiliation\\
  Address\\
  Country}
\email{author2@work}

\address{Murray Stokely\\
  Google, Inc.\\
  1600 Amphitheatre Parkway\\
  Mountain View, CA 94043\\
  USA}
\email{mstokely@google.com}