### 2014-01-03 papers/rjournal/JSSwrapper.tex
Add a JSSwrapper around the same content used for the RJournal article.  Edits should still be made to the same eddelbuettel-francois-stokely.Rnw file, and the Makefile still only builds the RJournal version by default, but now you can type pdflatex JSSwrapper instead and manually see what it would look like when formatted for JSS instead.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@687 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[article]{jss}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%
% Local helpers to make this more compatible with R Journal style.
%
\newcommand{\CRANpkg}[1]{\pkg{#1}}
\RequirePackage{fancyvrb}
\RequirePackage{alltt}
\DefineVerbatimEnvironment{example}{Verbatim}{}

%% almost as usual
\author{Dirk Eddelbuettel\\Debian and R Projects \And 
        Murray Stokely\\Google, Inc}
\title{\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in R}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Dirk Eddelbuettel, Murray Stokely} %% comma-separated
\Plaintitle{RProtoBuf: Efficient Cross-Language Data Serialization in R}
\Shorttitle{\pkg{RProtoBuf}: Protocol Buffers in R} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
Modern data collection and analysis pipelines often involve
a sophisticated mix of applications written in general purpose and
specialized programming languages.  Protocol Buffers are a popular
method of serializing structured data between applications---while remaining
independent of programming languages or operating system.  The
\CRANpkg{RProtoBuf} package provides a complete interface to this
library.
}
\Keywords{r, protocol buffers, serialization, cross-platform}
\Plainkeywords{r, protocol buffers, serialization, cross-platform} %% without formatting
%% at least one keyword must be supplied

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Murray Stokely\\
  Google, Inc.\\
  1600 Amphitheatre Parkway\\
  Mountain View, CA 94040\\
  USA\\
  E-mail: \email{mstokely@google.com}\\
  URL: \url{http://www.stokely.org/}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\include{eddelbuettel-francois-stokely}
%% include your article here, just as usual
%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.

%\section[About Java]{About \proglang{Java}}
%% Note: If there is markup in \(sub)section, then it has to be escape as above.

\end{document}